The Causes of Damage in Toronto Fire Incidents

### 2024-09-27 paper/paper.qmd
step4.1: Some Editing on paper

From (@fig-cause-compare-dollar) and (@fig-cause-compare-casualties),we see that among the top five causes of fires we identified,  each of them are substantially different in the damage they create.
Unattended ignition source is the most dangerous in causing death, but on average cause relatively less damage to properties. This suggests that causalities and property damage are different in nature and have to be treated differently. This observation also captures the nature of unattended unattended ignition sources, which often catch individuals off-guard, leading to a higher risk. 
The vast difference also highlights the importance of identifying the cause of fire, which may help us lower potential  property damage. And help us in understanding how to treat the fire.
The mean dollar loss for all causes of fire is 38,105.88, while the mean dollar loss for undetermined causes is significantly higher at 63,034. This indicates that fires of undetermined origin often result in more severe damage, possibly due to delayed detection and complex fire behavior. Undetermined fires are more likely to involve unusual circumstances, complicating both firefighting efforts and post-incident analysis. Understanding these patterns could be crucial for improving firefighting strategies.

@@

ggplot(top_fire_causes, aes(x = reorder(Possible_Cause, -Count), y = Avg_Civilian_Casualties)) +
  geom_bar(stat = "identity", fill = "orange", width = 0.6) +  # Adjust bar width
  geom_text(aes(label = round(Avg_Civilian_Casualties, 3)), size = 4, color = "black") +  # Add average casualties labels
  theme_minimal(base_size = 15) +  # Increase base font size for readability
  labs(title = "Average Civilian Casualties for Top 5 Causes",
       x = "Possible Cause",
       y = "Average Casualties") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6),  # Rotate x-axis labels
        axis.text.y = element_text(size = 6),  # Increase y-axis text size
        plot.margin = margin(t = 20, r = 20, b = 20, l = 20),  # Add margins
        panel.grid.major.x = element_blank())  # Remove grid lines
```
\newpage
## Number of responding personnel

@@

  (@fig-personnel-compare-casualties) is created from 25769 relevant observations, it draws out the relationship between number of responding personnel and civilian casualties. We can see as the number of personnel increases, the casualties increase, this behavior is expected since we would expect a more damaging fire to receive more attention. However, the average casualties reaches a peak when the personnel responding reaches 80-90, and decreases afterward.

@@

```{r}
#| label: fig-personnel-compare-casualties
#| fig-cap: Compare responding personnel with civilian casualties
#| include: true
#| echo: false
#| warning: false
#| message: false
#| console: false

@@

(@fig-personnel-compare-dollar-loss) is created from 23449 relevant observations, it draws out the relationship between number of responding personnel and dollar loss. We see that the damage of fire increases as number of personnel increase without exceptions. Which is our expected behavior and did not privide much insight comparing to (@fig-personnel-compare-casualties).

```{r}
#| label: fig-personnel-compare-dollar-loss
#| fig-cap: Compare dollar loss with civilian casualties
#| include: true
#| echo: false
#| warning: false
#| message: false

>>>>

From @fig-cause-compare-dollar and @fig-cause-compare-casualties,we see that among the top five causes of fires we identified,  each of them are substantially different in the damage they create.
Unattended ignition source is the most dangerous in causing death, but on average cause relatively less damage to properties. This suggests that causalities and property damage are different in nature and have to be treated differently. This observation also captures the nature of unattended unattended ignition sources, which often catch individuals off-guard, leading to a higher risk. 
The vast difference also highlights the importance of identifying the cause of fire, which may help us lower potential  property damage. And help us in understanding how to treat the fire.
The mean dollar loss for all causes of fire is 38,105.88, while the mean dollar loss for undetermined causes is significantly higher at 63,034. This indicates that fires of undetermined origin often result in more severe damage, possibly due to delayed detection and complex fire behavior. Undetermined fires are more likely to involve unusual circumstances, complicating both firefighting efforts and post-incident analysis. Understanding these patterns could be crucial for improving firefighting strategies.

@@

ggplot(top_fire_causes, aes(x = reorder(Possible_Cause, -Count), y = Avg_Civilian_Casualties)) +
  geom_bar(stat = "identity", fill = "orange", width = 0.6) +  # Adjust bar width
  geom_text(aes(label = round(Avg_Civilian_Casualties, 3)), size = 4, color = "black") +  # Add average casualties labels
  theme_minimal(base_size = 11) +  # Increase base font size for readability
  labs(title = "Average Civilian Casualties for Top 5 Causes",
       x = "Possible Cause",
       y = "Average Casualties") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6),  # Rotate x-axis labels
        axis.text.y = element_text(size = 5),  # Increase y-axis text size
        plot.margin = margin(t = 20, r = 20, b = 20, l = 20),  # Add margins
        panel.grid.major.x = element_blank())  # Remove grid lines
```
\newpage

@@

@fig-personcomparewithcasualties created from 25769 relevant observations, it draws out the relationship between number of responding personnel and civilian casualties. We can see as the number of personnel increases, the casualties increase, this behavior is expected since we would expect a more damaging fire to receive more attention. However, the average casualties reaches a peak when the personnel responding reaches 80-90, and decreases afterward.

@@

```{r}
#| label: fig-personcomparewithcasualties
#| fig-cap: Compare responding personnel with civilian casualties
#| include: true
#| echo: false
#| warning: false
#| message: false

@@

@fig-personcomparewithdollarloss is created from 23449 relevant observations, it draws out the relationship between number of responding personnel and dollar loss. We see that the damage of fire increases as number of personnel increase without exceptions. Which is our expected behavior and did not provide much insight comparing to @fig-personcomparewithcasualties.

```{r}
#| label: fig-personcomparewithdollarloss
#| fig-cap: Compare responding personnel with dollar loss
#| include: true
#| echo: false
#| warning: false
#| message: false
